---
title: "Climate data reporting analysis"
date: "September 29 2017"
output: html_document
---
```{r setup, include=F}
knitr::opts_chunk$set(comment = "#>")
```

## Analysis of climate reporting practices in published ecology studies
Morueta-Holme N, Oldfather MF, Olliff-Yang R, Weitz AP, Levine CR, Kling M, Riordan EC, Merow C, Sheth SN, Thornhill AH, Ackerly DD


### 0. Prepare data
Load libraries
```{r message=FALSE}
#rm(list=ls())
require(ggplot2)
require(dplyr)
require(gmodels)
require(tidyr)
```

Read in the data and select relevant columns
```{r, message=FALSE}
rawData1 = read.csv('Climate_reporting_dataset1.csv', as.is=T) 
rawData2 = read.csv('Climate_reporting_dataset2.csv', as.is=T)

# Select papers that are OK for analysis, and the relevant columns
df1 = rawData1[,c('ID','publication_year','journal','study_focus','geographic_extent_of_study','tense_description','climate_date_range','climate_data_source','citation_type','description_access_to_source','source_sufficient')]

df2 = rawData2[,c('ID','publication_year','journal','study_focus','geographic_extent_of_study','tense_description','climate_date_range','climate_data_source','citation_type','description_access_to_source','source_sufficient')]
                    
#merge the two tables
df = rbind(df1,df2)
rm(rawData1,rawData2,df1,df2)

#lump continental and global categories
df$geographic_extent_of_study[which(df$geographic_extent_of_study=='continental' | df$geographic_extent_of_study=='global')] = 'continental/global'


#Rename date range to 'yes/no'
df$climate_date_range[which(df$climate_date_range==TRUE)] = 'yes'
df$climate_date_range[which(df$climate_date_range==FALSE)] = 'no'


```


### 1. General summaries

What is the distribution of geographic extent of study across papers?
```{r include=T}
ExtentSum <-
  df %>%
  group_by(geographic_extent_of_study) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))
ExtentSum
```

What is the distribution of study focus across papers?
```{r include=T}
FocusSum <-
  df %>%
  group_by(study_focus) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))
FocusSum
```

How many papers provide climate date range?
```{r include=T}
DateRangeSum <-
  df %>%
  group_by(climate_date_range) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))
DateRangeSum
```

What are the types of climate data reported in the studies?
```{r, include=T}
DataSource <-
  df %>%
  group_by(climate_data_source) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

DataSource
```


### 2. Evalution of subset for detailed data source information
```{r, include=T}

df_sub <- df[df$citation_type!='',]
```

What type of citation is given?
```{r, include=T}
CitationSum <-
  df_sub %>%
  group_by(citation_type) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

CitationSum
```

Is there information on how to access the climate data?
```{r, include=T}
AccessSum <-
  df_sub %>%
  group_by(description_access_to_source) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

AccessSum
```

Is the citation given sufficient to retrieve the raw data?
```{r, include=T}
CitationSuffSum <-
  df_sub %>%
  group_by(source_sufficient) %>%
  summarise(n=n()) %>%
  mutate(rel.freq = round(100*n/sum(n), 0))

CitationSuffSum
```

If some sort of citation is given, is there also information on how to access the data?
```{r, include=F}
df_sub$some_citation <- 'no'
df_sub$some_citation[which(df_sub$citation_type!='incomplete'&df_sub$citation_type!='none')] <- 'yes'

t1 <- CrossTable(df_sub$some_citation, df_sub$description_access_to_source)

SourceAccess <- apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
```{r}
SourceAccess

```

### 3. Crosstabulations for evaluating reporting practices depending on study focus and geographic extent

#### Results across study focus
Climate data source given for papers with different study focus?
```{r include=F}
t1 <- CrossTable(df$study_focus, df$climate_data_source)
ResFocus1 <-apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
```{r}
ResFocus1
```
Climate date range given for papers with different study focus?
```{r include=F}
t1 <- CrossTable(df$study_focus, df$climate_date_range)
ResFocus2 <-apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
```{r}
ResFocus2
```
#### Results across geographic extent
```{r include=F, echo=F}
t1 <- CrossTable(df$geographic_extent_of_study, df$climate_data_source)
ResExtent1 <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

t1 <- CrossTable(df$geographic_extent_of_study, df$climate_date_range)
ResExtent2 <-apply(t1$prop.row,2,function(x) {round(x*100,1)})

t1 <- CrossTable(df$geographic_extent_of_study, df$study_focus)
ResExtent3 <- apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
Climate data source given for different geographic extents?
```{r}
ResExtent1
```
Climate date range given for different geographic extents?
```{r}
ResExtent2
```
Distribution of study foci across geographic extents?
```{r}
ResExtent3
```

### 4. Temporal analysis
```{r include=F}
# Lump years into decadal time periods except for 2015 papers
df$decade <- plyr::round_any(df$publication_year, 10, floor)
df$decade[df$publication_year>=2015] <- 2015

t1 <- CrossTable(df$decade, df$climate_data_source)
SourcePeriod <- apply(t1$prop.row,2,function(x) {round(x*100,1)})

t1 <- CrossTable(df$decade, df$climate_date_range)
RangePeriod <- apply(t1$prop.row,2,function(x) {round(x*100,1)})

t1 <- CrossTable(df$decade, df$study_focus)
FocusPeriod <- apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
Trends in reporting of type of climate data
```{r}
SourcePeriod
```
Trends in reporting of climate date range
```{r}
RangePeriod
```
Trends in focus of study
```{r}
FocusPeriod
```

### 5. Plot temporal trends
```{r echo=F}
# define time periods

#df$paper_focus <- factor(df$paper_focus, levels=sort(unique(df$paper_focus)))

# reporting per time period
dfs <- group_by(df, decade) %>%
      summarize(source=mean(climate_data_source=="not_specified"),
                dates=mean(climate_date_range=="no")) %>%
      gather(stat, rate, source, dates) %>%
      mutate(stat=factor(stat, levels=c("dates", "source"),
                         labels=c("no climate period", "no climate data type")))


p <- ggplot() +
      geom_bar(data=df,  aes(decade, fill=study_focus), position="fill") +
      geom_line(data=dfs, aes(decade, rate, linetype=stat), size=1) + 
      geom_point(data=dfs, aes(decade, rate, shape=stat), size=4) +
      scale_fill_manual(values=c("red3", "yellow3", "steelblue3")) +
      scale_shape_manual(values=c(16, 1)) +
      scale_y_continuous(labels=function(x) paste0(round(x*100,1), "%")) +
      scale_x_continuous(labels=function(x) paste0(x, "s"),
                         breaks=unique(dfs$decade)) +
      theme_minimal() +
      theme(legend.title=element_blank(), legend.position="right",
            legend.direction="vertical",
            panel.grid.minor=element_blank(), panel.grid.major.x=element_blank()) +
      labs(y="Percent of papers",
           x="Publication period")
p
#ggsave("trends.pdf", p, width=7, height=5, units="in")


```

### 6. Tense used to describe climate?
When reporting or not reporting climate date ranges, what tense is used for describing climate?
```{r include=F}
table(df$tense)
t1 <- CrossTable(df$climate_date_range, df$tense)
DateTense <- apply(t1$prop.row,2,function(x) {round(x*100,1)})
```
```{r}
DateTense
``` 
